DATA 621 - Business Analytics and Data Mining

### 2020-12-02 Final_project/Final Project Paper - Group 2.Rmd
Running on US only

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(doParallel)
library(knitr)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggplot2)
library(scales)
library(caret)
library(data.table)
```
```{r loadData}
dataHeader <- "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/"
ca_path <- paste0(dataHeader, "CAvideos.csv")
us_path <- paste0(dataHeader, "USvideos.csv")
gb_path <- paste0(dataHeader, "GBvideos.csv")
fr_path <- paste0(dataHeader, "FRvideos.csv")
de_path <- paste0(dataHeader, "DEvideos.csv")
droppedFields <- c('description', 'thumbnail_link')
ca_set <- fread(ca_path, drop = droppedFields)
us_set <- fread(us_path, drop = droppedFields)
gb_set <- fread(gb_path, drop = droppedFields)
fr_set <- fread(fr_path, drop = droppedFields)
de_set <- fread(de_path, drop = droppedFields)
nobsUS <- nrow(us_set)
nobsCA <- nrow(ca_set)
nobsGB <- nrow(gb_set)
nobsFR <- nrow(fr_set)
nobsDE <- nrow(de_set)
nobsTot <- nobsUS + nobsCA + nobsGB + nobsFR + nobsDE
```

@@

We manually scrubbed the data---which had numerous issues---so that we had five
clean sets of statistics for the United States, Canada, Great Britain, France,
and Germany. We identified both numeric and factor predictors, and engineered
features for convenience as well. We will use these features to investigate
relationships with actual views using standard linear and count (e.g. Poisson)
regression models.

@@

# Experimentation & Results
## Data Exploration
```{r ca_data_prep}
# Calculating the mean of the CA dataset
ca_mean <- mean(ca_set$views)

ca_set <- ca_set %>%
  # Add country label
  mutate(country = 'CA') %>% 
  # Calculate above/below view mean
  mutate(view_target = ifelse(views > ca_mean, 1L, 0L)) %>% 
  # Count number of tags
  mutate(tag_count = 1L + (str_count(ca_set$tags, ';'))) %>%
  # Now that we have the tag counts, we no longer need the tags themselves
  select(-tags)
```
```{r de_data_prep}
de_mean <- mean(de_set$views)

de_set <- de_set %>%
  mutate(country = 'DE') %>%
  mutate(view_target = ifelse(views > de_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(de_set$tags, ';'))) %>%
  select(-tags)
```
```{r fr_data_prep}
fr_mean <- mean(fr_set$views)

fr_set <- fr_set %>%
  mutate(country = 'FR') %>%
  mutate(view_target = ifelse(views > fr_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(fr_set$tags, ';'))) %>%
  select(-tags)
```
```{r gb_data_prep}
gb_mean <- mean(gb_set$views)

gb_set <- gb_set %>%
  mutate(country = 'GB') %>%
  mutate(view_target = ifelse(views > gb_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(gb_set$tags, ';'))) %>%
  select(-tags)
```
```{r us_data_prep}
us_mean <- mean(us_set$views)

us_set <- us_set %>%
  mutate(country = 'US') %>%
  mutate(view_target = ifelse(views > us_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(us_set$tags, ';'))) %>%
  select(-tags)
```
```{r merge_datasets}
# Since each country has the same format, we can just stack them one on top of
# the other. Much faster than merging.
full_set <- rbind(ca_set, de_set, fr_set, gb_set, us_set)

# Key the data table
setkey(full_set, title, country, trending_date)

# Calculate Global mean
full_mean <- mean(full_set$views)

# Calculate above/below global mean
full_set <- full_set %>%
  mutate(view_global_target = ifelse(views > full_mean, 1L, 0L))

# Convert trending_date to a date
full_set[, trending_date := as.IDate(trending_date, format = "%y.%d.%m")]

# Make name shorter for display purposes
setnames(full_set, c('comment_count', 'tag_count'), c('comments', 'tags'))
```
```{r mungeData}
# Categorize the variable names
boolVars <- c('comments_disabled', 'ratings_disabled', 'video_error_or_removed')
intVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
facVars <- c('country', 'category_id', 'view_target', 'view_global_target')
dateVars <- c('trending_date', 'publish_time')
charVars <- c('video_id', 'title', 'channel_title')

# Convert factors into factors. Set US as base factor for country
full_set[, `:=`(country = factor(country,
                                 levels = c('US', 'CA', 'GB', 'FR', 'DE'),
                                 labels = c('US', 'CA', 'GB', 'FR', 'DE')),
                category_id = as.factor(category_id),
                view_target = factor(view_target, levels = c(0, 1),
                                     labels = c(0, 1)),
                view_global_target = factor(view_global_target,
                                            levels = c(0, 1),
                                            labels = c(0, 1)))]
# Melt for numerics
fullSetN <- melt(full_set, id.vars = c(charVars, dateVars, boolVars, facVars),
                 variable.factor = FALSE, variable.name = 'measure',
                 value.name = 'value')

# Melt for factors except country
fullSetF <- melt(full_set,
                 measure.vars = c('category_id', 'view_target', 'view_global_target'),
                 variable.factor = FALSE, variable.name = 'measure',
                 value.name = 'value')
```
```{r summaryStats}
statsN <- data.table()
# Loop through countries and calculate statistics. Table shown in appendix.
for (i in unique(fullSetN$country)) {
  statsN <- rbind(statsN, cbind(country = i,
                  fullSetN[country == i,
                           .(Mean = mean(value, na.rm = TRUE),
                             SD = sd(value, na.rm = TRUE),
                             Min = min(value, na.rm = TRUE),
                             Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                             Median = median(value, na.rm = TRUE),
                             Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                             Max = max(value, na.rm = TRUE),
                             IQR = IQR(value, na.rm = TRUE)),
                           keyby = c('measure')]))
}
# Add global stats
statsN <- rbind(statsN,
                cbind(country = "Global",
                      fullSetN[,
                               .(Mean = mean(value, na.rm = TRUE),
                                 SD = sd(value, na.rm = TRUE),
                                 Min = min(value, na.rm = TRUE),
                                 Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                                 Median = median(value, na.rm = TRUE),
                                 Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                                 Max = max(value, na.rm = TRUE),
                                 IQR = IQR(value, na.rm = TRUE)),
                               keyby = c('measure')]))
```

@@

```{r numLogDensities, fig.height=4.5, fig.width=8}
ggplot(fullSetN[measure != 'views'], aes(x = value, color = measure)) +
  geom_density(kernel = "epanechnikov") +
  facet_wrap(~ country) +
  ggtitle("Predictor Density (semi-log scale)") +
  scale_x_log10(label = scientific)
```

### Factor Predictors
While tabular representation of factor predictors is difficult, a distribution
of factors by value may prove informative. It is clear from the graphs below
that interests vary by country. Category `24` seems to be the universal mode.
Except for Great Britain, where they love their category `10`!

```{r factorPlot1, fig.height=4.5, fig.width=8}
ggplot(fullSetF[measure == 'category_id'], aes(y = value)) +
  geom_bar() +
  facet_wrap(~ country) +
  ggtitle("Distribution by Category & Country") +
  scale_x_continuous(labels = comma)
```

```{r factorPlot2, fig.height=3, fig.width=8}
ggplot(fullSetF[measure != 'category_id'],
             aes(x = value, fill = measure)) +
  geom_bar(position = 'dodge2') +
  facet_wrap(~ country) +
  ggtitle("Country and Global Mean Views Exceedence") +
  scale_y_continuous(labels = comma)
```

As with many events which follow the Pareto rule, most of the views are 
attributable to a minority of the videos, as seen by the 0/1 disparity on both
an individual country and a global basis. Great Britain has the closest parity
on a global basis. British viewers are closest to the global average.

@@

```{r subsetExtraction}
# Select the row numbers of the first entry of the latest trending date by title
# and by country. There are 34 duplicates. Use this as our restricted data set
# corresponding to the most recent view count. The inner set of brackets gets
# the row number (called V1) and the outer set is a simple subset by those row
# numbers.
fullExtract <- full_set[
  full_set[, .I[which.max(trending_date)], by = c('title', 'country')]$V1
  ]
```

@@

```{r addFeatures}
fullExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
                   lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
```

## Model Building
We will first separate 20% of the data as a true holdout set. It is on this data
that our models will be compared.

```{r holdOut}
set.seed(617)
seenIDX <- createDataPartition(fullExtract$views, p = 0.8)$Resample1
seenSet <- fullExtract[seenIDX, ]
hideSet <- fullExtract[-seenIDX, ]
```

We will train models on the remaining 80% of the data.

### Model 1: Linear Regression
Keeping country as a predictor is of less interest. It will most likely track
some measure of population. It is of more interest to see if there are
differences in the factors between countries. To capture this by interactions
would be combinatorially unfeasible. Instead, five models will be run, each on
a country specific subset. This is akin to a manual implementation of a rules
based regression model

```{r modelLinear1, cache = 2L}
# Using stepAIC means no cross-validation. Train on entire dataset.
trC <- trainControl(method = 'none', allowParallel = TRUE)
cl <- makePSOCKcluster(5)
registerDoParallel(cl)
set.seed(181)
for (ct in unique(fullExtract$country)) {
  linMod <- train(views ~ likes + dislikes + comments + tags + category_id +
                    comments_disabled + ratings_disabled +
                    video_error_or_removed + perfect + lkratio,
                  data = fullExtract[country == ct],
                  family = gaussian(link = 'identity'), direction = 'both',
                  method = 'glmStepAIC', trace = 0, trControl = trC)
  assign(paste0('linMod', ct), linMod)
}
stopCluster(cl)
lM1US <- linModUS$finalModel
lM1CA <- linModCA$finalModel
lM1GB <- linModGB$finalModel
lM1FR <- linModFR$finalModel
lM1DE <- linModDE$finalModel
clM1US <- linModUS$finalModel$coefficients
clM1CA <- linModCA$finalModel$coefficients
clM1GB <- linModGB$finalModel$coefficients
clM1FR <- linModFR$finalModel$coefficients
clM1DE <- linModDE$finalModel$coefficients
nclM1US <- names(clM1US)
nclM1CA <- names(clM1CA)
nclM1GB <- names(clM1GB)
nclM1FR <- names(clM1FR)
nclM1DE <- names(clM1DE)
nnclM1US <- length(nclM1US)
nnclM1CA <- length(nclM1CA)
nnclM1GB <- length(nclM1GB)
nnclM1FR <- length(nclM1FR)
nnclM1DE <- length(nclM1DE)
length(nclM1US) <- 22
length(nclM1GB) <- 22
length(nclM1FR) <- 22
length(nclM1DE) <- 22
```

\footnotesize
```{r m1Table}
kable(data.frame(cbind(nclM1CA, nclM1DE, nclM1US, nclM1FR, nclM1GB)),
      caption = "Model Coefficients by Country")
```
\normalsize

The most parsimonious model is the one for Great Britain with only `r nnclM1GB`
parameters. The most complex is the one for Canada with `r nnclM1CA`.

@@

```{r loadLibraries}
```
```{r loadData}
```
<!--  Data Exploration -->
```{r ca_data_prep}
```
```{r de_data_prep}
```
```{r fr_data_prep}
```
```{r gb_data_prep}
```
```{r us_data_prep}
```
```{r merge_datasets}
```
```{r mungeData}
```
```{r summaryStats}
```
```{r factorPlot1}
```
```{r factorPlot2}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(jsonlite)
library(knitr)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggplot2)
library(scales)
library(caret)
library(data.table)
```
```{r loadData}
currentPath <- getwd()
dataPath <- "Data2"
us_path <- file.path(currentPath, dataPath, "USvideos.csv")

colClass <- c(rep('character', 4L), 'integer', 'POSIXct', 'character',
              rep('double', 4L), rep('boolean', 3L))
us_set <- fread(us_path, encoding = 'UTF-8', colClasses = colClass)
nobsUS <- nrow(us_set)

# Category IDs
us_cat_path <- file.path(currentPath, dataPath, "US_category_id.json")
us_cats <- fromJSON(us_cat_path)

us_cats <- data.table(id = as.integer(us_cats$items$id),
                      category = us_cats$items$snippet[, 2])
setkey(us_cats, id)
```

@@

We manually scrubbed the data and discovered that the country-specific files
really were not! They neither refer to videos created by country nor do they
refer to views *specific* to country. Rather they are the total number of views
and other predictors for that video on that day as collected by someone within
that country. Meaning that aggregation is almost always multiplying. There may
be some videos unique to a specific country---one which was not viewed in other
countries, but they are many magnitudes smaller than those seen by all. As the
United States had the most observations, we decided to analyze its data set.

With the data, we identified both numeric and factor predictors, and engineered
features for convenience as well. We will use these features to investigate
relationships with actual views using linear regression models.

@@

# Experimentation & Results
## Data Exploration

```{r dataPrep}
# Substitute category for numeric ID through joining
us_set <- us_cats[us_set, on = 'id == category_id']

# Change data types for convenience, add target bucket, count tags and then
# get rid of actual tags and category IDs

us_mean <- mean(us_set$views)

us_set[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
              views = as.double(views),
              tag_count = 1L + str_count(us_set$tags, '[|]'),
              view_target = factor(ifelse(views > us_mean, 1L, 0L),
                                   levels = c(0, 1), labels = c(0, 1)),
              id = NULL,
              category = as.factor(category))
       ][, `:=`(tags = NULL,
                category = relevel(category, 'Entertainment'))]

# Make name shorter for display purposes
setnames(us_set, c('comment_count', 'tag_count'), c('comments', 'tags'))

# Categorize the variable names
boolVars <- c('comments_disabled', 'ratings_disabled', 'video_error_or_removed')
numVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
facVars <- c('category', 'view_target')
dateVars <- c('trending_date', 'publish_time')
charVars <- c('video_id', 'title', 'channel_title')

# Melt for numerics
usSetN <- melt(us_set, measure.vars = numVars, variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

# Melt for factors
usSetF <- melt(us_set,
               measure.vars = c('category', 'view_target'),
               variable.factor = FALSE, variable.name = 'measure',
               value.name = 'value')

# Table of summary statistics
statsN <- usSetN[, .(Mean = mean(value, na.rm = TRUE),
                     SD = sd(value, na.rm = TRUE),
                     Min = min(value, na.rm = TRUE),
                     Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                     Median = median(value, na.rm = TRUE),
                     Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                     Max = max(value, na.rm = TRUE),
                     IQR = IQR(value, na.rm = TRUE)),
                 keyby = c('measure')]
```

@@

```{r numLogDensities, fig.height=4.5, fig.width=8}
ggplot(usSetN[measure != 'views'], aes(x = value, color = measure)) +
  geom_density(kernel = "epanechnikov") +
  ggtitle("Predictor Density (semi-log scale)") +
  scale_x_log10(label = scientific)
```

### Factor Predictors
While tabular representation of factor predictors is difficult, a distribution
of factors by value may prove informative. It is clear from the graphs below
that interests vary by category. `Entertainment` seems to be the most common in
the US with `Music` coming in second.

```{r factorPlot1, fig.height=4.5, fig.width=8}
ggplot(usSetF[measure == 'category'], aes(y = value)) +
  geom_bar() +
  ggtitle("Distribution by Category") +
  scale_x_continuous(labels = comma)
```

```{r factorPlot2, fig.height=3, fig.width=8}
ggplot(usSetF[measure != 'category'], aes(x = value, fill = measure)) +
  geom_bar(position = 'dodge2') +
  ggtitle("Country and Global Mean Views Exceedence") +
  scale_y_continuous(labels = comma)
```

As with many events which follow the Pareto rule, most of the views are 
attributable to a minority of the videos, as seen by the 0/1 disparity.

@@

```{r subsetExtraction}
# Select the row numbers of the first entry of the latest trending date by title
# and by country. There are 34 duplicates. Use this as our restricted data set
# corresponding to the most recent view count. The inner set of brackets gets
# the row number (called V1) and the outer set is a simple subset by those row
# numbers.
usExtract <- us_set[
  us_set[, .I[which.max(views)], by = c('title')]$V1
  ]
```

@@

```{r addFeatures}
usExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
                 lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
```

## Model Building
We will first separate 20% of the data as a true holdout set. It is on this data
that our models will be compared. We will train models on the remaining 80% of
the data.

```{r holdOut}
set.seed(617)
seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
seenSet <- usExtract[seenIDX, ]
hideSet <- usExtract[-seenIDX, ]
```

### Multiple Linear Regression
Linear regression may be the best known algorithm used when analyzing a
continuous numeric outcome. It searches for a linear relationship of the
predictors that minimizes the squared error between the "predictor" function and
the observations [@Sheather2009AMA]. This model will engage in stepwise feature
using the AIC as the optimization metric.

```{r modelLinear1}
# Using stepAIC means no cross-validation. Train on entire dataset.
trC <- trainControl(method = 'none')

# Capture the unique country codes
set.seed(181)
linMod1 <- train(views ~ likes + dislikes + comments + tags + category +
                   comments_disabled + ratings_disabled +
                   video_error_or_removed + perfect + lkratio,
                 data = usExtract, family = gaussian(link = 'identity'),
                 method = 'glmStepAIC', direction = 'both', trace = 0,
                 trControl = trC)
```

\footnotesize
```{r m1Table}
kable(summary(linMod1$finalModel)$coefficients,
      caption = "Model 1 Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

The sign of these coefficients make sense in the main. The most predictive
feature for views is `likes` followed by `dislikes`. Of course, one usually
views a video at least once prior to rating it. We set the baseline category to
`Entertainment`, considering it was the most popular. Therfore we expected
negative coefficients for other categories. Interestingly, `Film & Animation`
is positively correlated with likes over and above `Entertainment`, despite
having fewer views.

Other observations are that the magnitude of the coefficients is similar to that
of the intercept, indicating that the predictors provide large swings to final
views. The "worst" predictor by magnitude is the `Nonprofits & Activism`
category; its negative coefficient is more than twice the baseline intercept
views. The best, interestingly is if ratings are disabled. This is
counterintuitive at first, but is logical. The amount of `likes`, `dislikes`,
and `comments` are many orders of magnitude greater than 1. Therefore, their
coefficients can be much smaller. A Boolean variable is either 1 or 0, therefore
its coefficient is much greater even if its actual contribution is lower.

Variable importance can be seen from the table below. The further down the list
a coefficient preceded by a **`-`** is, the more important it is and the more it
affected the AIC.
\footnotesize
```
                                  Df   Deviance    AIC
<none>                               7.1243e+16 212209
+ `categoryAutos & Vehicles`       1 7.1222e+16 212209
+ tags                             1 7.1223e+16 212209
- `categoryPeople & Blogs`         1 7.1278e+16 212210
- categoryEducation                1 7.1278e+16 212210
- perfectTRUE                      1 7.1279e+16 212210
+ video_error_or_removedTRUE       1 7.1237e+16 212210
+ `categoryPets & Animals`         1 7.1239e+16 212210
+ categoryGaming                   1 7.1239e+16 212210
+ `categoryTravel & Events`        1 7.1240e+16 212210
+ comments_disabledTRUE            1 7.1240e+16 212210
+ categorySports                   1 7.1240e+16 212210
+ `categoryScience & Technology`   1 7.1241e+16 212211
+ categoryShows                    1 7.1243e+16 212211
- `categoryHowto & Style`          1 7.1287e+16 212211
- `categoryFilm & Animation`       1 7.1299e+16 212212
- `categoryNews & Politics`        1 7.1321e+16 212214
- categoryMusic                    1 7.1338e+16 212215
- `categoryNonprofits & Activism`  1 7.1345e+16 212216
- ratings_disabledTRUE             1 7.1463e+16 212227
- categoryComedy                   1 7.1467e+16 212227
- lkratio                          1 7.3555e+16 212413
- dislikes                         1 8.9091e+16 213650
- comments                         1 8.9807e+16 213701
- likes                            1 2.0217e+17 218939
```
\normalsize

@@

```{r loadLibraries}
```
```{r loadData}
```
```{r dataPrep}
```
```{r numLogDensities}
```
```{r factorPlot1}
```
```{r factorPlot2}
```
